---
title: "p8105_hw1_ls4236"
author: "Liliang Su"
date: "2025-09-11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1

Load the `moderndive` library and `early_january_weather` dataset.
```{r data_loading}
library(moderndive)

data("early_january_weather")
```

### Data Description

First of all, I try to write a brief description of the dataset using inline R code.

```{r check_the_data}
head(early_january_weather)
```
As the above table shows, in `early_january_weather` dataset, there are `r nrow(early_january_weather)` rows (samples) and `r ncol(early_january_weather)` columns (variables), including `r colnames(early_january_weather)`. Specifically, key variables include:

- `origin`(`character`): the location of weather station.
- `year, month, day, hour`(`integer`): Time of recording.
- `temp` (`double`): air temperature in Fahrenheit.
- `humid` (`double`): relative humidity between 0 and 1.
- `wind_dir, wind_speed` (`double`): wind direction (in degrees) and speed(in mph).
- `time_hour` (`POSIXct`): Date and hour of the recording as a POSIXct date.

Notably, the `temp` column has an average temperature of `r mean(early_january_weather$temp, na.rm = TRUE)`.

### Plotting

Secondly, I try to create a scatterplot of `temp`(y) vs `time_hour`(x), with color points using the `humid` variable.

```{r}
plot_temp_th <- ggplot(early_january_weather, aes(x = time_hour,  y = temp, color = humid)) + geom_point() + labs(x = "Time (hourly)", y = "Temperature", color = "Humidity")

plot_temp_th

ggsave("plot_temp_th.png", plot_temp_th, width = 8, height = 5)
```
Overall, the scatterplot shows that the temperature has increased over time, rising from about 30°F in early January to around 50°F in mid-January. Despite the upward trend, there has been frequent drastic fluctuations in daily temperature. Humidity has also shown a generally positive correlation with temperature, with higher values more common at warmer temperatures in mid-January.


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
